---
output: html_document
params:
  file_path_zip: "/Users/lucassilbernagel/Documents/_Master HWR Berlin_/3nd Semester/Internship Swiss TPH/First week/03a-TIMCI-SPA-fa.zip"
  file_name_csv: "07-TIMCI-fa - audit.csv"
  display_code_chunks: TRUE
  set_title: "some_default"
title: "`r params$set_title`"
author: "Lucas Silbernagel"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---
`r if(!params$display_code_chunks) {"\\begin{comment}"}`
```{r, eval=TRUE, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(readr)
library(ggplot2)
library(openxlsx)
library(knitr)
library(tibble)
library(stringr)
```

## I. Loading the data
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# Unzip and extract ODK data from ODK zip
extract_data_from_odk_zip <- function(odk_zip, csv_name, start_date = NULL, end_date = NULL) {
  
  t <- tempdir()
  utils::unzip(odk_zip, exdir = t)
  fs::dir_ls(t)
  raw_odk_data <- readr::with_edition(1, readr::read_csv(file.path(t, csv_name)))
}

df <- as.data.frame(extract_data_from_odk_zip(params$file_path_zip, params$file_name_csv))

# Formatting dates from integer (in ms) to timestamp
df$start = as.POSIXct(df$start/1000, origin="1970-01-01")
df$end = as.POSIXct(df$end/1000, origin="1970-01-01")
summary(df)
```

## II. Deriving new features
**Time spent per event**
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# subtracting end from start date
df$time_spent = as.numeric((df$end - df$start)/60)
```

**Question**
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# function to split the node strings so that only the question name remains 
create_question <- function(str){
  key=tail(unlist(strsplit(str,"/")),1)
  return(key)
}

df$question = sapply(df$node, create_question)
```

**QuestionII with distinct values for step type**
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# list to decode the response values for step type
step_type_decoding = c("registration", "waiting", "triage", "consultation", 'pharmacy', 'pharmacy', 'walking', 'combined registration', 'other')

# replacing the response encoder value with 9 to match the list order
df$`new-value`[df$`new-value`==99] <- 9

# initializing a second question column where step type is differentiated into its several options
df$questionII = sapply(df$node, create_question)
df$questionII[df$questionII=='a1_a_4'] = 'scanning QR'
df$questionII[df$questionII=='a1_a_4a'] = 'enter ID manually'
df$questionII[df$questionII=='b1_4'] = 'select district'
df$questionII[df$questionII=='fcode'] = 'select facility'
for (row in 1:nrow(df)){
  if (df$questionII[row]=='step_type' & !is.na(df$questionII[row])){
    df$questionII[row]<- step_type_decoding[as.integer(df$`new-value`[row])]
  }
}
```

**Time until a response was changed + changes stream for step type events**
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
df <- df %>%
# bringing the data in the right order   
  arrange(`instance ID`, node, start) %>%
# adding two empty columns to store the new features in
  add_column(time_till_change=NA) %>%
  add_column(changed_from=NA)

# iterating over the df and computing the time it took until an answer was changed + adding what the question was before 
for (i in 1:nrow(df)){
  if (df$`old-value`[i]==df$`new-value`[i-1] && !is.na(df$`old-value`[i]) && !is.na(df$`new-value`[i-1]) ){
    df$time_till_change[i] <- as.numeric(df$start[i]-df$end[i-1])
    df$changed_from[i] <- df$questionII[i-1] 
  } else{
    next
  }
}
head(df)
```
`r if(!params$display_code_chunks) {"\\end{comment}"}`

`r if(params$display_code_chunks) {"\\begin{comment}"}`
```{r, include=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(readr)
library(ggplot2)
library(openxlsx)
library(knitr)
library(tibble)
library(stringr)
```

```{r, include=FALSE, message=FALSE, warning=FALSE}
# Unzip and extract ODK data from ODK zip
extract_data_from_odk_zip <- function(odk_zip, csv_name, start_date = NULL, end_date = NULL) {
  
  t <- tempdir()
  utils::unzip(odk_zip, exdir = t)
  fs::dir_ls(t)
  raw_odk_data <- readr::with_edition(1, readr::read_csv(file.path(t, csv_name)))
}

df <- as.data.frame(extract_data_from_odk_zip(params$file_path_zip, params$file_name_csv))

# Formatting dates from integer (in ms) to timestamp
df$start = as.POSIXct(df$start/1000, origin="1970-01-01")
df$end = as.POSIXct(df$end/1000, origin="1970-01-01")
summary(df)
```

```{r, include=FALSE, message=FALSE, warning=FALSE}
# Time spent per event

# subtracting end from start date
df$time_spent = as.numeric((df$end - df$start)/60)
```


```{r, include=FALSE, message=FALSE, warning=FALSE}
#Question

# function to split the node strings so that only the question name remains 
create_question <- function(str){
  key=tail(unlist(strsplit(str,"/")),1)
  return(key)
}

df$question = sapply(df$node, create_question)
```


```{r, include=FALSE, message=FALSE, warning=FALSE}
# QuestionII with distinct values for step type

# list to decode the response values for step type
step_type_decoding = c("registration", "waiting", "triage", "consultation", 'pharmacy', 'pharmacy', 'walking', 'combined registration', 'other')

# replacing the response encoder value with 9 to match the list order
df$`new-value`[df$`new-value`==99] <- 9

# initializing a second question column where step type is differentiated into its several options
df$questionII = sapply(df$node, create_question)
df$questionII[df$questionII=='a1_a_4'] = 'scanning QR'
df$questionII[df$questionII=='a1_a_4a'] = 'enter ID manually'
df$questionII[df$questionII=='b1_4'] = 'select district'
df$questionII[df$questionII=='fcode'] = 'select facility'
for (row in 1:nrow(df)){
  if (df$questionII[row]=='step_type' & !is.na(df$questionII[row])){
    df$questionII[row]<- step_type_decoding[as.integer(df$`new-value`[row])]
  }
}
```

```{r, include=FALSE, message=FALSE, warning=FALSE}
# Time until a response was changed + changes stream for step type events

df <- df %>%
# bringing the data in the right order   
  arrange(`instance ID`, node, start) %>%
# adding two empty columns to store the new features in
  add_column(time_till_change=NA) %>%
  add_column(changed_from=NA)

# iterating over the df and computing the time it took until an answer was changed + adding what the question was before 
for (i in 1:nrow(df)){
  if (df$`old-value`[i]==df$`new-value`[i-1] && !is.na(df$`old-value`[i]) && !is.na(df$`new-value`[i-1]) ){
    df$time_till_change[i] <- as.numeric(df$start[i]-df$end[i-1])
    df$changed_from[i] <- df$questionII[i-1] 
  } else{
    next
  }
}
head(df)
```
`r if(params$display_code_chunks) {"\\end{comment}"}`

```{r, include = FALSE, message=FALSE, warning=FALSE}
if (params$display_code_chunks){
  knitr::opts_chunk$set(
    echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE
  )
} else{
  knitr::opts_chunk$set(
    echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE
  )
}
```

## Loading the data
```{r}
# Unzip and extract ODK data from ODK zip
extract_data_from_odk_zip <- function(odk_zip, csv_name, start_date = NULL, end_date = NULL) {
  
  t <- tempdir()
  utils::unzip(odk_zip, exdir = t)
  fs::dir_ls(t)
  raw_odk_data <- readr::with_edition(1, readr::read_csv(file.path(t, csv_name)))
}

df <- as.data.frame(extract_data_from_odk_zip(params$file_path_zip, params$file_name_csv))

# Formatting dates from integer (in ms) to timestamp
df$start = as.POSIXct(df$start/1000, origin="1970-01-01")
df$end = as.POSIXct(df$end/1000, origin="1970-01-01")
summary(df)
```

## Deriving new features
**Time spent per event**
```{r}
# subtracting end from start date
df$time_spent = as.numeric((df$end - df$start)/60)
```

**Question**
```{r}
# function to split the node strings so that only the question name remains 
create_question <- function(str){
  key=tail(unlist(strsplit(str,"/")),1)
  return(key)
}

df$question = sapply(df$node, create_question)
```

**QuestionII with distinct values for step type**
```{r}
# list to decode the response values for step type
step_type_decoding = c("registration", "waiting", "triage", "consultation", 'pharmacy', 'pharmacy', 'walking', 'combined registration', 'other')

# replacing the response encoder value with 9 to match the list order
df$`new-value`[df$`new-value`==99] <- 9

# initializing a second question column where step type is differentiated into its several options
df$questionII = sapply(df$node, create_question)
df$questionII[df$questionII=='a1_a_4'] = 'scanning QR'
df$questionII[df$questionII=='a1_a_4a'] = 'enter ID manually'
df$questionII[df$questionII=='b1_4'] = 'select district'
df$questionII[df$questionII=='fcode'] = 'select facility'
for (row in 1:nrow(df)){
  if (df$questionII[row]=='step_type' & !is.na(df$questionII[row])){
    df$questionII[row]<- step_type_decoding[as.integer(df$`new-value`[row])]
  }
}
```

**Time until a response was changed + changes stream for step type events**
```{r}
df <- df %>%
# bringing the data in the right order   
  arrange(`instance ID`, node, start) %>%
# adding two empty columns to store the new features in
  add_column(time_till_change=NA) %>%
  add_column(changed_from=NA)

# iterating over the df and computing the time it took until an answer was changed + adding what the question was before 
for (i in 1:nrow(df)){
  if (df$`old-value`[i]==df$`new-value`[i-1] && !is.na(df$`old-value`[i]) && !is.na(df$`new-value`[i-1]) ){
    df$time_till_change[i] <- as.numeric(df$start[i]-df$end[i-1])
    df$changed_from[i] <- df$questionII[i-1] 
  } else{
    next
  }
}
head(df)
```
**Checking for outliers by lookig at the distribution of time_spent.**
```{r}
#plotting cdf to see distribution of the time_spent feature 
ecdf = ecdf(df$time_spent[!is.na(df$time_spent)])
plot(ecdf, main="CDF of Time Spent per Event")

# plotting a histogram just to reassure
hist(df$time_spent[!is.na(df$time_spent)])
```

**Very few events made up for very long event times. To avoid biases in summary statistics, we remove the largest 5 % of values**
```{r}
df_clean = subset(df, time_spent<quantile(df$time_spent,0.95, na.rm=TRUE))
```

## 2. Median time spent per question
### 2.1 Table of all questions and their median time
```{r}
df_median_time_per_question <- df_clean %>%
  filter(event=='question') %>%
  group_by(question) %>%
  summarise(median_time = median(time_spent)) %>%
  arrange(desc(median_time))
kable(df_median_time_per_question)

df_median_time_per_questionII <- df_clean %>%
  filter(event=='question') %>%
  group_by(questionII) %>%
  summarise(median_time = median(time_spent)) %>%
  arrange(desc(median_time))
kable(df_median_time_per_questionII)
```

## 3. Input changes per question
### 3.1 Table with counts of input changes per question and per question II
```{r}
df_changes_per_question <- df %>%
  filter(event=='question', !is.na(`old-value`), !is.na(`new-value`)) %>%
  group_by(question) %>%
  summarise(count=n()) %>%
  arrange(desc(count))
kable(df_changes_per_question)

df_changes_per_questionII <- df %>%
  filter(event=='question', !is.na(`old-value`), !is.na(`new-value`)) %>%
  group_by(questionII) %>%
  summarise(count=n()) %>%
  arrange(desc(count))
kable(df_changes_per_questionII)
```

### 3.2. Table with median time until input is changed per question
```{r}
df_time_till_change <- df %>%
  filter(!is.na(time_till_change)) %>%
  group_by(question) %>%
  summarise(median_time_till_change = median(time_till_change)) %>%
  arrange(desc(median_time_till_change))
kable(df_time_till_change)
```

### 3.3 Table with count of old-new value pairs
```{r}
df_stream <- df %>%
  filter(!is.na(changed_from)) %>%
  group_by(questionII, changed_from) %>%
  summarise(count=n())%>%
  arrange(desc(count))

kable(df_stream)
```

### 3.4 Table showing time till change outliers (for all data without anyhting removed)
```{r}
subset(df, time_till_change>quantile(df$time_till_change, 0.90, na.rm=TRUE)) %>% 
  arrange(desc(time_till_change)) %>%
  select(questionII, time_spent, time_till_change, start, end)
```

## 4. Metrics per instance
### 4.1 Distribution of duration per instance 
```{r}
df_duration_per_inst <- df %>%
  group_by(`instance ID`) %>%
  summarise(duration_per_inst = max(end, na.rm=T) - min(start, na.rm=T)) %>%
  mutate(duration_per_inst_mins =as.numeric(duration_per_inst/60^2)) %>%
  arrange(desc(duration_per_inst))
df_duration_per_inst

hist(as.numeric(df_duration_per_inst$duration_per_inst_mins), breaks=20, main="Histogram of Duration per Instance in minutes", xlab="Duration in minutes")
```
